---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SMMAL

<!-- badges: start -->
<!-- badges: end -->

SMMAL is an R package for estimating the Average Treatment Effect (ATE) using semi-supervised learning (SSL), tailored for settings with limited treatment/outcome labels but rich covariates and surrogate variables. It enhances efficiency and robustness over supervised methods by leveraging unlabeled data and supports high-dimensional models via cross-fitting, flexible model fitting, and adaptive LASSO.


## Installation

A github version can be found at this link: https://github.com/ShuhengKong/SMMAL

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(SMMAL)

# Load the example dataset included with the package
file_path <- system.file("extdata", "sample_data.rds", package = "SMMAL")
dat <- readRDS(file_path)

temp <- data.frame(dat$X)
temp[,] <- NA
# Estimate ATE using the SMMAL pipeline
output <- SMMAL(
  Y = dat$Y,
  A = dat$A,
  S = data.frame(dat$S),
  X = data.frame(dat$X),
  nfold = 5,
  cf_model = "bspline"
)

# View the results
print(output)

```

## CALDERA input files
| Column        | Description |
| ----------    | ---------- |
| Y    | Observed outcomes. Can be continuous or binary  |
| A       | Treatment indicator. Must be binary |
| S   | Surrogates |
| X        | Covariates |
| nfold    | Number of cross-validation folds. Default is 5. |
| cf_model    | The modeling method to use in cross-fitting. Default is "bspline". Other values are "xgboost","randomforest" |


## CALDERA output
| Column        | Description |
| ----------    | ---------- |
| est        | estimated value of ATE |
| se    | standard error of ATE |

