---
title: "SMMAL_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SMMAL_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

This vignette demonstrates how to use the SMMAL package to estimate the Average Treatment Effect (ATE) using semi-supervised machine learning. We provide an example dataset and walk through the required input format and function usage.


## Import Sample data. 

Sample data contain 1000 observations with 60% of Y and A missing at random. 
Y is the outcome. 
A is the treatment indicator. 
X are the covariates. 
S are the surrogates.

For the sample data, missingness occurs at random and is encoded as NA.
This package can handle datasets with a high proportion of missing values, but it requires a sufficiently large sample size to ensure that each fold in cross-validation contains at least 20 labeled observations.

```{r}
library(SMMAL)

file_path <- system.file("extdata", "sample_data_withmissing.rds", package = "SMMAL")
dat <- readRDS(file_path)
```

## Prepare Inputs

Input file S and X needs to be data frame, even if they are vectors.


```{r}
  # Y and A are numeric vector 
  Y <- dat$Y
  A <- dat$A
  
  # S and X needs to be data frame
  S <- data.frame(dat$S)
  X <- data.frame(dat$X)
``` 


## Estimate ATE with SMMAL & Output

Users can choose which model to use for the nuisance functions by setting the cf_model parameter.
If no cf_model is indicated, the default value is "bspline".

After cross-validation and prediction, the best-performing model is selected based on the lowest cross-entropy (log loss).
Users can control how many folds are used in cross-validation by setting the nfold parameter.
If no nfold is indicated, the default value is 5.
  
```{r}
 SMMAL_output1 <- SMMAL(Y=Y,A=A,S=S,X=X)
 print(SMMAL_output1)
```

Other options for cf_model are "xgboost" 
  
```{r}
SMMAL_output2 <- SMMAL(Y=Y,A=A,S=S,X=X,cf_model= "xgboost")
print(SMMAL_output2)
```

or "random forest"

```{r}
SMMAL_output3 <- SMMAL(Y=Y,A=A,S=S,X=X,cf_model= "randomforest")
print(SMMAL_output3)
```


## Using Your Own custom_model_fun

Users may customize the feature‐selection or penalization strategy by supplying their own function through the custom_model_fun argument. To do so, pass a function that meets these requirements:

1. Function Signature
It must accept exactly these arguments (in this order):
X, Y, foldid_labelled, sub_set, labeled_indices, nfold, log_loss

(X, Y, foldid_labelled, sub_set, labeled_indices, and nfold are used internally by SMMAL to partition and fit the data.)

(log_loss is a function for computing cross‐entropy (log‐loss). Your function should call log_loss(true_labels, predicted_probs) to evaluate each tuning parameter.)

2. Return Value
It must return a list of length equal to the number of “ridge” penalty values defined in param_fun(). Each element of that list should be a numeric vector of length n containing out‐of‐fold predicted probabilities for all observations—i.e., it should stack together predictions from every held‐out fold (no NA values, except where Y is genuinely missing).

Below is an example showing how to plug in the packaged ada_lasso() as custom_model_fun. In practice, you could substitute any function with the same signature and return type:

```{r}
 SMMAL_output4 <- SMMAL(Y=Y,A=A,S=S,X=X, custom_model_fun = ada_lasso)
 print(SMMAL_output4)
```




