---
title: "SMMAL_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SMMAL_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Introduction

This vignette demonstrates how to use the SMMAL package to estimate the Average Treatment Effect (ATE) using semi-supervised machine learning. We provide an example dataset and walk through the required input format and function usage.


Import Sample data. 

Sample data contain 1000 observations with 60% of Y and A missing at random. 
Y is the outcome. A is the treatment indicator. X are the covariates. S are the surrogates.
For the sample data, missingness occurs at random and is encoded as NA.
This package can handle datasets with a high proportion of missing values, but it requires a sufficiently large sample size to ensure that each fold in cross-validation contains at least 20 labeled observations.

```{r}
library(SMMAL)

file_path <- system.file("extdata", "sample_data_withmissing.rds", package = "SMMAL")
dat <- readRDS(file_path)
```


Input file S and X needs to be data frame, even if they are vectors.


```{r}
  # Y and A are numeric vector 
  Y <- dat$Y
  A <- dat$A
  
  # S and X needs to be data frame
  S <- data.frame(dat$S)
  X <- data.frame(dat$X)
``` 


Estimate ATE with SMMAL & Output

Users can choose which model to use for the nuisance functions by setting the cf_model parameter.
If no cf_model is indicated, the default value is "bspline".

After cross-validation and prediction, the best-performing model is selected based on the lowest cross-entropy (log loss).
Users can control how many folds are used in cross-validation by setting the nfold parameter.
If no nfold is indicated, the default value is 5.
  
```{r}
SMMAL_output1 <- SMMAL(Y=Y,A=A,S=S,X=X)
print(SMMAL_output1)
```

Other options for cf_model are "xgboost" 
  
```{r}
SMMAL_output2 <- SMMAL(Y=Y,A=A,S=S,X=X,cf_model= "xgboost")
print(SMMAL_output2)
```

or "randomforest"

```{r}
SMMAL_output3 <- SMMAL(Y=Y,A=A,S=S,X=X,cf_model= "randomforest")
print(SMMAL_output3)
```


Users may also replce some of the built in function by their own choice. 
Take the function ada_lasso as an example.


The ada_lasso function takes X (covariates) and Y (outcome) as input and selects the most predictive covariates for Y.
If X is a vector, a logistic regression (glm) is used to determine whether the covariate should be retained.
If no covariates are selected, the function returns NULL.
Downstream functions in the package are designed to detect a NULL return and will halt execution with an error message indicating that X must be a non-empty data frame.


Below, we demonstrate the output of the ada_lasso() function.

When data frame X has multiple columns.

```{r}
SMMAL_output4 <- ada_lasso(Y=Y,X=X)
str(SMMAL_output4)
```

The ada_lasso() function returns a reduced version of the input covariate matrix X, containing only the variables that are predictive of the outcome Y, as determined by adaptive LASSO or a univariate logistic regression test.

Below, we demonstrate the output when no no covariates was selected.

```{r}
SMMAL_output5 <- ada_lasso(Y=Y,X=X[, 1:2, drop = FALSE])
print(SMMAL_output5)
```

In this case, the output is a single NULL value, indicating that no covariates were selected.



When data frame X has only multiple column.
The ada_lasso() function uses a univariate logistic regression test when the input X is a single-column data frame (i.e., a vector of covariates). 
If the test finds that the covariate is not statistically significant (p-value ≥ 0.05), the function determines that the covariate should not be retained.

```{r}
SMMAL_output6 <- ada_lasso(Y=Y,X=X[, 8, drop = FALSE])
str(SMMAL_output6)
```

In this case, the output is the covariate itself.


```{r}
SMMAL_output7 <- ada_lasso(Y=Y,X=X[, 1, drop = FALSE])
print(SMMAL_output7)
```

In this case, the output is a single NULL value, indicating that no covariate was selected.


